\subsubsection{\stid{6.01} LANL ATDM Tools}

\paragraph{Overview}
The Kitsune Project, part of LANL's ATDM CSE efforts, provides a
compiler-focused infrastructure for improving various aspects of the
Exascale programming environment.  At present our most efforts are
focus on advanced LLVM compiler and tool infrastructure supporting the
use of a \emph{parallel-aware} intermediate representation.  In
addition, we are actively involved in the Flang Fortran compiler that
is now an official sub-project within the overall LLVM infrastructure.
All these efforts include include interactions across ECP as well as
within the broader LLVM community and industry.  

\paragraph{Key Challenges}
A key challenge to our effort is the broad adoption and agreement that
a parallel intermediate representation is beneficial and needed within
LLVM.  This not only requires showing benefits but also provided a
full implementation for evaluation and feedback from the community.
In addition, significant new compiler capabilities represents a
significant effort and involves many complexities and technical
challenges.  These efforts and the process of up-streaming potential
design and implementation changes do involve some amount of time and
associated risk. 

Additional challenges come by a range of complex issues surrounding
target architectures for exascale systems.  The use of the LLVM
infrastructure helps reduce many challenges here as many processor
vendors and system providers now leverage and use LLVM for their
commercial compilers.

\paragraph{Solution Strategy}
Given the project challenges, our approach considers aspects of
today's node-level programming systems (e.g. OpenMP and Kokkos) and
popular programming languages (e.g. C++ and Fortran) into
consideration and aims to improve and expand upon their capabilities
to address the needs of ECP.  This allows us to attempt to strike a
balance between incremental improvements to existing infrastructure
along with more aggressive techniques that seeks to provide innovative
solutions to help both manage risk and the ability to introduce new
breakthrough technologies.

Unlike current designs, our approach introduces the notion of explicit
parallel constructs into the LLVM intermediate representation, building
off of work done at MIT on Tapir~\cite{2.3.6.01:kitsune:Schardl:2017}. 
We are exploring extensions to this work as well as making some changes
to fundamental data structures within the LLVM infrastructure to assist
and improve in better analysis and optimization passes. 

\paragraph{Recent Progress}

Our primary focus is the delivery of capabilities for LANL's ATDM
Ristra application (AD 2.2.5.01).  In support of the requirements for
Ristra we are targeting the lowering of Kokkos constructs directly
into the parallel-IR representation.  At present this requires
explicit lowering of Kokkos constructs and we have basic support
for \texttt{parallel\_for} and \texttt{parallel\_reduce} in place.  In
addition we are looking at replacement of LLVM's dominator tree, a key
data structure for optimizations, including parallelization, and
memory usage analysis, with a \emph{dominator directed-acyclic-graph}
(DAG).  This work is currently underway and should near its initial
implementation early in the 2020 calendar year.  In addition, we are
actively watching recent events within the LLVM community around
multi-level intermediate representations (MLIR) and the relationship
they have with parallel semantics, analysis, optimization, and code
generation.  Furthermore, there are still discussions within the
community about general details behind parallel intermediate forms. 

\paragraph{Next Steps}

The key next steps of our feature set is to complete use cases of
Kokkos that match the needs for Ristra.  Where possible we will also
explore the broader set of use cases within ECP's overall use of
Kokkos constructs.  The Kitsune toolchain is still very much an
active \emph{proof-of-concept} effort based on Clang (C and C++) with
plans to add support for Fortran via the newly established Flang front
end within LLVM (ST 2.3.2.12).  Even though it is not yet production
ready we are actively updating and releasing source code and the supporting
infrastructure for deployment as an early evaluation candidate.  In
addition to these components we will actively begin to explore targeting
the exascale systems (Aurora, Frontier, and El Capitan).

