ECP ST Product Dictionary
Note: This page is still under construction.

The ECP Software Technology (ST) Product Dictionary is the official list of publicly recognized 
names to which ECP ST efforts contribute.  While ST teams use an expanded product 
namespace, the list on this page indicates the eventual access point for ST product development 
efforts.

This table lists only those products that are typically recognizable to users. Examples:
1.	MPI is commonly known by users. MPICH and OpenMPI both provide implementations 
of that product.
2.	Fortran is a product. Flang is a particular Fortran product. LLVM is a backend for some 
Fortran compilers.
3.	FFT is a product. FFTX, FFT-ECP provide FFT capabilities through interchangeable 
interfaces. 
4.	C++ is a product. Clacc provides capabilities for Clang, as does LLVM.

Product 
Dictionary 
List
ST products that deliver capabilities through public products 
(comma separated list)

ADIOS

AML

Ascent
Cinema, ParaView, Visit, VTK-m
BLAS
SLATE
C
LLVM
C++
LLVM
Caliper

Catalyst

CHAI

Cinema

CUDA

Darshan

DTK

Dyninst

E4S

FFT
FFTX, FFT\_ECP
FleCSI

Flux

Fortran
LLVM/Flang
GASNet

Ginkgo

HDF5

HPCToolkit

hypre

Kokkos

KokkosKernels

LAPACK

Legion

libEnsemble

MarFS

MFEM

MPI
MPICH, Open MPI
OpenACC
Clacc/LLVM
OpenCL
-
OpenMP
SOLLVE/LLVM
PAPI

Papyrus

Paraview

PaRSEC

PETSc/TAO

PnetCDF

PowerStack

RAJA

MPI-IO
ROMIO
ScaLAPACK
SLATE
SCR

SICM

Spack

SPOT

STRUMPACK

SUNDIALS

SuperLU

SYCL

SZ

TASMANIAN

TAU

Trilinos

UMap

Umpire

Unify

UPC++

VeloC

Visit

VTK-m

xSDK

ZFP


