\subsubsection{LLNL ATDM Data \& Viz Projects: Workflow}

%%{\itshape
%%
%%	\begin{enumerate}
%%	\item Rename this file to your project WBS-projectname.tex, for example 2.3.3.01-XSDK4ECP.tex.
%%	\item Complete this template for your project.  Limit your text to two pages, not counting citations.  
%%	\item Please avoid changing the content of main.tex.  
%%	\item Put any references in a .bib file with the same root name, for example 2.3.3.01-XSDK4ECP.bib.
%%	\item Remember to include any image files you reference in your text.
%%    \item The files 2.3.3.01-XSDK4ECP.tex, 2.3.3.01-XSDK4ECP.bib and xSDK-diagram.jpeg are included as examples for your reference.  You can remove them from what you upload.
%%	\end{enumerate}
%%}

\paragraph{Overview} 

The ATDM Workflow project at LLNL has a long term vision to enable reproducible
and pedigreed computational science and introduce advanced analytics and machine
learning capabilities to our user community. We are building an
ecosystem of re-usable components for user workflows that will enhance the
end-to-end productivity of ASC HPC assets and enable users to introduce modern
data analytics technologies to their workflows. The Workflow project is focused
on three areas prioritized by the LLNL user community: problem setup, simulation
management, and data management and analytics. The project focuses on an
ecosystem of tools and libraries because there is no single overarching
workflow system that can meet the needs of users in the wide variety of domains
at LLNL. The tools produced by this project are focused on reducing extrinsic
cognitive load, that is, to reduce the bookkeeping and repetitive one-off
scripting that is required with current file-oriented workflows. The project
seeks to leverage modern data analytics capabilities, by providing tools that
users can leverage to inject simulation information into these systems. 

\paragraph{Key Challenges}

The three biggest challenges for implementing improved problem setup,
simulation management, and data management and analytics for users are: user
community heterogeneity and complexity, security requirements, and user
adoption. First, the large number of user domains and the number of largely
manually driven workflows in use precluded the installation of a single
encompassing workflow management and product lifecycle management capability.
Second, the security posture required to install tools at NNSA labs makes it
difficult to stand up federated infrastructures and servers.  Finally, adoption
of tools into existing workflows requires that the tools provide a clear value
proposition to the user community, and any tool that requires large changes in
existing practice faces an increased chance of being rejected by users. This is
particularly true for systems that seek to capture full provenance and pedigree
data: they often require centralized servers and that all workflow activities
take place in a particular tool or environment.

\paragraph{Solution Strategy}

We established three sub-projects to focus on aspects of the three areas prioritized by users:
\begin{enumerate}
\item \textbf{C2C:} the Contours to Codes project focused on building tools for
interchanging and streamlining the process of going from documents describing
experiments to the corresponding simulation configurations. C2C has adopted a
format for 2D geometry defined by Los Alamos National Laboratory, and has built
Python and C++ parsers for this toolset with features needed for LLNL
applications.  
\item \textbf{Siboka:} an ecosystem of scientific workflow components to
facilitate workflow and data management, increase pedigree/reproducibility, and
enable the use of modern analytics technologies.  \textbf{Siboka} includes a
set of Python packages and command line tools for capturing non-bulk data and
meta-data from simulation runs, importing data to SQL and Cassandra (a column
store), allow querying of data in those backends, and exporting results for use
in downstream tools. The core tool being developed is called \textbf{Sina},
which enables these  operations. In addition, a specification for a JSON
schema and toolset called \textbf{Mnoda} is being developed to generalize an
application specific tool created by the project to gather and query non-bulk
data and meta-data from simulation run directories.
\item \textbf{Syflux (formerly VV4ALE3D):} a prototype web portal for setting
up, running, and managing simulations and their results.  The original
prototype, VV4ALE3D, is integrated with the Lorenz web infrastructure deployed
in Livermore Computing.  The VV4ALE3D application is still deployed and hosted
on the Livemore Computing web infrastructure.  Syflux represents an updated
version based on the Django web infrastructure and hosted on project-managed
web servers outside the data center authentication realm. The team seeks to
develop into a problem setup data portal as well as a way to deploy workflow
management tools.  
\end{enumerate}
These projects represent composable tools targeting distinct aspect of user
pain points.  C2C enables interchange of problem setup information across
applications, and a format in which to store and curate the results of the time
consuming process of setting up new simulations.  Siboka presents the users
with distinct and focused Python packages that they can include in existing
scripts, so that they can incrementally adopt the technologies they see as
relevant to their work.  Syflux and associated exploration of Jupyter,
represents customizable ways to provide access via web-browser to simulation
data, problem setup and related databases, and documents.  In this way, the
team mitigates risk by providing the user community with a set of components
that encompasses best practices, common operations, and next-generation
capabilities, allowing the user community to leverage them to build custom
workflows where needed while reducing duplicated work.

\paragraph{Recent Progress}

textbf{C2C} recently released an initial C++ parser to LLNL application codes,
a key deliverable to enable adoption of a common 2D geometry format across our
user-base. The team has also defined an assembly file specification to allow
manipulation and composing of lower level geometric information, and
implemented a set of python tools for this. These tools are now being tested by
a few initial users and they have demonstrated that the tools can be used for
code inter-comparison.

The Siboka team has recently created an importer for \textit{Cassandra}, a
scalable, open-source column store used extensively by companies like Netflix
and Walmart. The team demonstrated the ability to
query and export data from SQL using the \textbf{Sina} package, and plans to
extend that functionality to the Cassandra back-end. The team is now exploring
how \textit{JupyterHub}, \textit{Jupyter Notebooks}, and the newly released (in
beta) \textit{JupyterLab} enhance sharing and reproducibilty while
incorporating these modern database capabilities.  The first version of the
\textbf{Mnoda} schema has been defined, and the Cassandra importer leverages
that tool to map simulation meta-data and non-bulk scalars and file paths to
the Cassandra and SQL back-ends.

The Syflux effort has stood up a Django-based web portal, and has enabled users
to upload and curate data and media files and implemented a basic permission
model for them.  The ability to cross-link resources (input files, output
files, and documents/reports) was also added to the prototype.

\paragraph{Next Steps}
\begin{enumerate}
\item \textbf{C2C:} the highest priority for the team is to integrate the C++
	library into the first application code and continue to enhanceme 
	assembly files. 
\item \textbf{Siboka:} Release an initial Mnoda schema, and create a specification for a C linkable API for writing data for applications that do not want to write the JSON output directly. Enhance the support for Cassandra and release an updated version of \textbf{Sina} to our users, as well a set of JupyterHub examples.
\item \textbf{Syflux}: prototype an integration of multiple databases, such as
	the Granta database into the portal.

\end{enumerate}
