\subsubsection{\stid{3.15} Sake Sub-project: Kokkos Kernels} \label{subsubsect:kokkoskernels}
\paragraph{Overview} 
The Kokkos Kernels~\footnote{https://github.com/kokkos/kokkos-kernels} subproject primarily
focuses on performance portable kernels for sparse/dense linear algebra, graphs, and machine
learning, with emphasis on kernels that are key to the performance of
several ECP applications. We work closely with ECP applications to identify the
performance-critical kernels and develop portable algorithms for these kernels.
The primary focus of this subproject is to support ECP application needs, develop new
kernels needed with an emphasis towards software releases, tutorials, boot camps
and user support. The Kokkos Kernels project also works closely with several vendors
(AMD, ARM, HPE/Cray, Intel, and NVIDIA) as part of both the ECP collaborations and
NNSA's Center for Excellence efforts. These collaborations will enable vendor solutions
that are targeted towards ECP application needs.

\paragraph{Key  Challenges}
There are several challenges in allowing ECP applications move to the hardware architectures
announced in the next few years. We highlight the four primary challenges here:
\begin{enumerate}
\item 
The next three supercomputers that will be deployed will have
three different accelerators from AMD, Intel and NVIDIA. While we have been expecting diversity of architectures, three
different architectures in such a short time frame adds pressure on the portability
solutions such as Kokkos Kernels to optimize and support the kernels on all the platforms.
\item
The design of several ECP applications and a software stack that rely on a component-based
approach results in an extremely high number of kernels launches on the accelerators, which
results in the latency costs becoming the primary bottleneck in scaling the applications.
\item
The change in the needs of applications from device-level kernels to smaller team-level kernels. Vendor
libraries are not ready for such a drastic change in software design.
\item 
The reliance of ECP applications on certain kernels that do not port well to the accelerator architectures.
\end{enumerate}

These challenges require a collaborative effort to explore new algorithmic choices,
working with the vendors to incorporate ECP needs into their library plans, to develop
portable kernels from scratch, and to deploy them in a robust software ecosystem. The Kokkos
Kernels project will pursue all of these choices in an effort to address these challenges.

\paragraph{Solution Strategy}
To mitigate these challenges, the team proposes the following actions.


    \subparagraph{Codesign Portable Kernels with Vendors and Applications}
    We rely on codesign of Kokkos Kernels implementations for
    specializations that are key to the performance of ECP applications. This
    requires tuning kernels even up to the problem sizes that are of interest
    to our users. Once we have developed a version, we provide these
    to all the vendors so their teams can optimize these kernels even
    further in vendor-provided math libraries.
   \subparagraph{Emphasis on Software Support and Usability}
	The Kokkos Kernels project devotes a considerable amount of time working with
	ECP applications, integrating the kernels into application codes, tuning
	for application needs, and providing tutorials and user support. We invest
	in delivering a robust software ecosystem that serves the
	needs of diverse ECP applications on all platforms of interest.
   \subparagraph{Invest in Algorithmic Research to Reduce Latency Costs and New Accelerator-Focused Approaches} To resolve latency cost issues, the Kokkos Kernels team is considering several solutions from computer science perspectives and also from algorithmic applied mathematics perspectives. 
   For example, from a computer science perspective, we are focusing on the
   use of streams or other latency reducing techniques such as cuda graphs. From
   the applied mathematics perspective we are developing new algorithms such
   as cluster-based approaches for preconditioners, such as Gauss-Seidel preconditioners,
   to reduce the number of kernel launches. 


\paragraph{Recent Progress}
\begin{itemize}
\item New batched linear solver algorithms are being developed in support of ECP applications (Pele, XGC).
Specifically, a preconditioned batched GMRES is now available and provides good performance on multiple GPU
architectures of interest to ECP.
\item Support for block sparse row matrices (BsrMatrix) is now available and associated matrix-vector
operation is available. Additional performance improvements are expected for the BsrMatrix mat-vec.
\item Further progress on the SYCL backend has been made and are available in the develop branch of
our repository. These changes will be included in release 3.6.0 and will result in further performance
portability for the libraries and applications we support (Trilinos, PETSc, ExaWind).
\item Kokkos Kernels team is currently updating documentation and wiki pages to keep up with release 3.6.0
new features. Adjustments to tutorial are being made when necessary to ensure the new capabilities are
advertised appropriately.
\end{itemize}

\paragraph{Preliminary Experiences on Early Access systems}
Kokkos Kernels engages all ECP facilities to ensure the library development is taking into account
the specificities of all the architectures applications will encounter. Some of the recent developments
include the development of new performance algorithms that leverage tensor core capabilities of
Nvidia V100 on Summit and A100 on Perlmutter. New support for rocBLAS and rocSPARSE has been deployed
and tunning for our Kernels on the HIP backend has been performed. These have been tested on EAS Spock
and Crusher at OLCF. Finally regular meetings are held with Argonne and Intel staff to ensure that the
library is deployed on JLSE Arcticus and provides adequate capabilities to users. These meetings have
resulted in significant improvements for our SYCL backend support.

\paragraph{Next Steps}

The Kokkos Kernels team has identified the following efforts for the next phase of the project.

	\subparagraph{A Major Software Release} Kokkos ecosystem 4.0 release 
	will be available to the ECP applications in FY 2022. This includes several new 
	kernels that are requested by ECP applications, performance improvements of
	kernels that are already being used by ECP applications and further support for
        new SYCL and OpenMP Target to enable applications to be tested on Intel architectures.
	\subparagraph{Develop New Sparse Batched and Block-Sparse Kernels} The team continues
        to work on sparse batched linear algebra kernels and associated dense batched kernels.
        Additionally, more support for BsrMatrix is forthcoming with Matrix-Matrix multiplication,
        Jacobi and Gauss-Seidel algorithms.
	\subparagraph{Collaborate with Vendors and Facilities} The Kokkos Kernels' team is
        continually working with vendors and facilities to improve the software environment
        on available supercomputers and early access systems.

