\subsubsection{\stid{5.04} Sandia ATDM Software Ecosystem and Delivery -
  Technology Demonstrator} 

\paragraph{Overview} 

This project is part of the NNSA/ASC program and is primarily focused
on the integration of programming models technologies in existing
applications. This project spans both the Code Development and
Applications (CDA) and Architecture and Software Development (ASD)
components of Sandia's ATDM program element, and this description only
covers the ASD portion of the project.

The purpose of this project is to broaden the impact of
programming model technologies developed in ATDM by integrating these
technologies into an existing ASC code suite, evaluating the ability
of these technologies to be incrementally added into a legacy code
base and to deliver performance improvements on current- and
next-generation computing platforms. In particular, this subproject
focuses on using and evaluating Kokkos and asynchronous many task
(AMT) programming models in the NGP Contact code. Kokkos is an
abstraction layer that implements a parallel programming model in C++
for writing performance portable applications targeting all major HPC
platforms. NGP Contact is a performance-critical proximity search
component relevant to many NNSA applications, including ASC production
applications that invoke mechanical or thermal contact, coupled
physics transfers, and particle algorithms. This project is important
the overall ECP efforts because it evaluates how well technologies
developed for next-generation applications can be leveraged by
existing large application code bases, such as the ASC Sierra toolkit,
in an effort to maximize the impact of ECP technologies. 

\paragraph{Key  Challenges}

The main challenge associated with this project is the incremental
integration of a new programming model and programming system into an
existing very large production application. This challenge is similar
in many ways to the incremental integration of OpenMP into an existing
large MPI application. As Kokkos is a higher-level abstraction layer
that can be implemented using OpenMP, there are additional challenges
around the ability of Kokkos' abstractions and interfaces to integrate
into an existing application and minimizing the performance penalty of
the abstraction layer for different underlying hardware
implementations.

\paragraph{Solution Strategy}

The strategy for this project is to produce demonstration applications
that drive the development of an AMT scheduling toolset and to enable
leveraging of Sierra-developed technologies to support ATDM
application milestones.

\paragraph{Recent Progress}

The Technology Demonstrator team has continued progress toward a 
performant, optimized GPU contact implementation. 
Two final parts of the contact algorithm are being ported to run on the GPU: 
interaction determination and interaction enforcement. The Sierra Solid 
Mechanics and NimbleSM teams have begun, and are continuing, joint work on 
porting the existing NimbleSM CPU-only routines and expect to test GPU 
execution. Initial performance data indicated that providing a full 
implementation of GPU contact could provide significant performance 
improvements over CPU-only execution, and performance data will be 
gathered once the routines are ready.

\paragraph{Next Steps}

The team's next focus will be to add a frictional contact capability to 
NimbleSM using node-face contact. This effort will include demonstrating 
correctness through a few canonical verification problems, optimization
 as needed, and evaluations of relative performance on hardware 
representative of LLNL's ATS-2 platform.

